---
title: "Sensitivity Analysis of Treatment Effect Modifications"
author: "Automated Report"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "../../")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(gt)
library(scales)
```

# Overview

This report summarises calibration performance following modification of treatment effects for statins and antihypertensives.

The report includes:

- Summary calibration metrics (ICI, E50, E90)
- Heatmaps of calibration performance
- Heatmaps showing changes relative to the original model
- Overlay calibration curves comparing scenarios
- Difference curves showing deviation from the original calibration curve

```{r scenario-definition}
scenario_adjustments <- expand.grid(
  statins = c(0.8, 0.9, 1, 1.1, 1.2),
  ah = c(0.8, 0.9, 1, 1.1, 1.2)
) |>
  filter(!(statins == 1 & ah == 1))

rows_1_unchanged <- which(
  scenario_adjustments$statins == 1 |
    scenario_adjustments$ah == 1
)

scenario_adjustments <- scenario_adjustments[rows_1_unchanged, ]
```

# Load Sensitivity Analysis Results

```{r load-results}
load_results <- function(gender){

  map_dfr(
    #seq_len(nrow(scenario_adjustments)),
    1:6,
    function(i){

      statins <- scenario_adjustments$statins[i]
      ah <- scenario_adjustments$ah[i]

      ici <- readRDS(
        paste0(
          "data/sens_treatment_effect_modified_calib_ph_ICI_",
          gender,
          "_statins", statins,
          "_ah", ah,
          ".rds"
        )
      )

      e50 <- readRDS(
        paste0(
          "data/sens_treatment_effect_modified_calib_ph_E50_",
          gender,
          "_statins", statins,
          "_ah", ah,
          ".rds"
        )
      )

      e90 <- readRDS(
        paste0(
          "data/sens_treatment_effect_modified_calib_ph_E90_",
          gender,
          "_statins", statins,
          "_ah", ah,
          ".rds"
        )
      )

      boot <- readRDS(
        paste0(
          "data/sens_treatment_effect_modified_calib_ph_CI_ICI_E50_E90_",
          gender,
          "_statins", statins,
          "_ah", ah,
          ".rds"
        )
      )

      tibble(
        gender = gender,
        statins = statins,
        ah = ah,
        ICI = ici,
        E50 = e50,
        E90 = e90,
        ICI_lwr = quantile(boot[,1], 0.025),
        ICI_upr = quantile(boot[,1], 0.975),
        E50_lwr = quantile(boot[,2], 0.025),
        E50_upr = quantile(boot[,2], 0.975),
        E90_lwr = quantile(boot[,3], 0.025),
        E90_upr = quantile(boot[,3], 0.975)
      )
    }
  )
}

female_results <- load_results(2)
male_results <- load_results(1)

results_all <- bind_rows(female_results, male_results)
```

# Load Original Model Results

```{r}
load_reference <- function(gender){

  tibble(
    gender = gender,
    ICI = readRDS(paste0("data/calib_ph_ICI_", gender, "_model4.rds")),
    E50 = readRDS(paste0("data/calib_ph_E50_", gender, "_model4.rds")),
    E90 = readRDS(paste0("data/calib_ph_E90_", gender, "_model4.rds"))
  )

}

reference_results <- bind_rows(
  load_reference(1),
  load_reference(2)
)

results_diff <-
  results_all |>
  left_join(
    reference_results,
    by = "gender",
    suffix = c("", "_ref")
  ) |>
  mutate(
    delta_ICI = ICI - ICI_ref,
    delta_E50 = E50 - E50_ref,
    delta_E90 = E90 - E90_ref
  )
```

# Summary Tables

```{r}
female_results |> arrange(ICI) |> gt()
```

```{r}
male_results |> arrange(ICI) |> gt()
```

# Heatmaps

```{r}
ggplot(female_results,
       aes(factor(statins), factor(ah), fill = ICI)) +
  geom_tile() +
  geom_text(aes(label = round(ICI, 4))) +
  labs(title = "Female ICI")
```

```{r}
ggplot(male_results,
       aes(factor(statins), factor(ah), fill = ICI)) +
  geom_tile() +
  geom_text(aes(label = round(ICI, 4))) +
  labs(title = "Male ICI")
```

# Change Relative to Original Model

```{r}
plot_delta_heatmap <- function(dat, metric, title){

  ggplot(
    dat,
    aes(
      factor(statins),
      factor(ah),
      fill = .data[[metric]]
    )
  ) +
    geom_tile() +
    geom_text(
      aes(label = round(.data[[metric]], 4))
    ) +
    scale_fill_gradient2(
      midpoint = 0,
      low = "blue",
      mid = "white",
      high = "red"
    ) +
    labs(
      title = title,
      x = "Statin multiplier",
      y = "AH multiplier"
    )

}

plot_delta_heatmap(
  filter(results_diff, gender == 2),
  "delta_ICI",
  "Female ΔICI"
)
```

```{r}
plot_delta_heatmap(
  filter(results_diff, gender == 1),
  "delta_ICI",
  "Male ΔICI"
)
```

```{r}
plot_delta_heatmap(
  filter(results_diff, gender == 2),
  "delta_E50",
  "Female ΔE50"
)
```

```{r}
plot_delta_heatmap(
  filter(results_diff, gender == 1),
  "delta_E50",
  "Male ΔE50"
)
```

```{r}
plot_delta_heatmap(
  filter(results_diff, gender == 2),
  "delta_E90",
  "Female ΔE90"
)
```

```{r}
plot_delta_heatmap(
  filter(results_diff, gender == 1),
  "delta_E90",
  "Male ΔE90"
)
```

# Calibration Curve Utilities

```{r}
load_smooth_curve <- function(gender, statins, ah){

  readRDS(
    paste0(
      "data/sens_treatment_effect_modified_calib_ph_df_smooth_",
      gender,
      "_statins", statins,
      "_ah", ah,
      ".rds"
    )
  )

}

load_reference_curve <- function(gender){

  readRDS(
    paste0(
      "data/calib_ph_df_smooth_",
      gender,
      "_model4.rds"
    )
  )

}
```

# Overlay Calibration Curves

The plots below compare the original calibration curve with selected treatment-effect modifications.

```{r}
plot_overlay <- function(gender){

  reference <- load_reference_curve(gender) |>
    mutate(scenario = "Original")

  statins_low <- load_smooth_curve(gender, 0.8, 1) |>
    mutate(scenario = "Statins 0.8")

  statins_high <- load_smooth_curve(gender, 1.2, 1) |>
    mutate(scenario = "Statins 1.2")

  ah_low <- load_smooth_curve(gender, 1, 0.8) |>
    mutate(scenario = "AH 0.8")

  ah_high <- load_smooth_curve(gender, 1, 1.2) |>
    mutate(scenario = "AH 1.2")

  curves <- bind_rows(
    reference,
    statins_low,
    statins_high,
    ah_low,
    ah_high
  )

  ggplot(curves,
         aes(pred, pred.obs, colour = scenario)) +
    geom_line(linewidth = 1) +
    geom_abline(
      slope = 1,
      intercept = 0,
      linetype = 2
    ) +
    labs(
      x = "Predicted risk",
      y = "Observed risk"
    )

}
```

## Female Overlay Curves

```{r}
plot_overlay(2)
```

## Male Overlay Curves

```{r}
plot_overlay(1)
```

# Difference Curves

Difference curves show:

Observed risk (scenario) - Observed risk (original)

Values above zero indicate higher observed risk than the original calibration curve at that level of predicted risk.

```{r}
plot_difference_curves <- function(gender){

  reference <- load_reference_curve(gender)

  scenarios <- list(
    "Statins 0.8" = c(0.8, 1),
    "Statins 1.2" = c(1.2, 1),
    "AH 0.8" = c(1, 0.8),
    "AH 1.2" = c(1, 1.2)
  )

  diff_curves <- lapply(
    names(scenarios),
    function(name){

      vals <- scenarios[[name]]

      curve <- load_smooth_curve(
        gender,
        vals[1],
        vals[2]
      )

      curve |>
        mutate(
          difference =
            pred.obs - reference$pred.obs,
          scenario = name
        )
    }
  )

  diff_curves <- bind_rows(diff_curves)

  ggplot(
    diff_curves,
    aes(pred, difference, colour = scenario)
  ) +
    geom_hline(
      yintercept = 0,
      linetype = 2
    ) +
    geom_line(linewidth = 1) +
    labs(
      x = "Predicted risk",
      y = "Difference from original curve"
    )

}
```

## Female Difference Curves

```{r}
plot_difference_curves(2)
```

## Male Difference Curves

```{r}
plot_difference_curves(1)
```

# Interpretation Template

This section is intentionally written as a template rather than a conclusion.

Suggested points to review:

1. Which scenarios produce the largest increase in ICI?
2. Which scenarios produce the largest increase in E50 and E90?
3. Are the effects similar between males and females?
4. Does calibration appear more sensitive to changes in statin effects or antihypertensive effects?
5. Do deviations occur uniformly across the risk spectrum or mainly among high-risk individuals?
6. Do the overlay plots indicate a systematic shift in calibration or a change in curve shape?
7. Do the difference curves suggest clinically meaningful distortion of predicted risks?

No automated interpretation is included so that conclusions can be drawn directly from the reported results.
